---
title: "VZ" 
author: "Author: Daniela Cassol (danielac@ucr.edu) and Thomas Girke (thomas.girke@ucr.edu)"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`" 
output:
  BiocStyle::html_document:
    toc_float: true
    code_folding: show
  BiocStyle::pdf_document: default
package: systemPipeR
vignette: |
  %\VignetteIndexEntry{RIBO-Seq Workflow Template}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
fontsize: 14pt
bibliography: bibtex.bib
---

```{css, echo=FALSE}
pre code {
white-space: pre !important;
overflow-x: scroll !important;
word-break: keep-all !important;
word-wrap: initial !important;
}
```

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options(width=60, max.print=1000)
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE")), 
    tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r setup, echo=FALSE, messages=FALSE, warnings=FALSE, eval=FALSE}
suppressPackageStartupMessages({
    library(systemPipeR)
})
```

# VZ fct

add the table
| Function Name     | Description                                                                                                           | Category        |
|-------------------|-----------------------------------------------------------------------------------------------------------------------|-----------------|
| `genWorkenvir`    | Generates workflow templates provided by systemPipeRdata helper package / or from the individuals’ pipelines packages | Accessory       |
| `loadWorkflow`    | Constructs `SYSargs2` object from CWL param and targets files                                                         | ``SYSargs2``    |
| `renderWF`        | Populate all the command-line in an `SYSargs2` object                                                                 | ``SYSargs2``    |
| `subsetWF`        | Subsetting `SYSargs2` class slots                                                                                     | ``SYSargs2``    |
| `runCommandline`  | Executes command-line software on samples and parameters specified in `SYSargs2` object                               | ``SYSargs2``    |
| `clusterRun`      | Runs command-line software in parallel mode on a computer cluster                                                     | ``SYSargs2``    |
| `writeTargetsout` | Write updated targets out to file/Generate targets file with reference                                                | ``SYSargs2``    |
| `output_update`   | Updates the output files paths in the `SYSargs2` object                                                               | ``SYSargs2``    |
| `singleYML`       | Create automatically the `param.yml`                                                                                  | ``SYSargs2``    |
| `createWF`        | Create automatically `param.cwl` and the `param.yml` based on the command line                                        | ``SYSargs2``    |
| `config.param`    | Custom configuration of the CWL param files from R                                                                    | ``SYSargs2``    |
| `initWF`          | Constructs `SYSargsList` workflow control module (S4 object) from script file                                         | ``SYSargsList`` |
| `configWF`        | Control of which step of the workflow will be run and the generation of the new RMarkdown                             | ``SYSargsList`` |
| `runWF`           | Runs all the R chunk define in the RMarkdown file or a subset, e.g. `runWF[1:3]`                                      | ``SYSargsList`` |
| `renderReport`    | Render Scientific Report based on RMarkdown                                                                           | ``SYSargsList`` |
| `subsetRmd`       | Write updated subset Rmarkdown of R chunk with text associate in the step selected                                    | ``SYSargsList`` |
| `plotWF`          | Plot visual workflow designs and topologies with different graphical layouts                                          | ``SYSargsList`` |
| `statusWF`        | Return the overview of the workflow steps computational status                                                        | ``SYSargsList`` |
| `evalCode`        | Turn eval option `TRUE` or `FALSE` on RMarkdown file                                                                  | Accessory       |
| `tryCL`           | Checks if third-party software or utility is installed and set in the PATH                                            | Accessory       |

## barplot
describe to the user how to use this fct

```{r}
add the example run from the help files
```
## heatmap

# How to use this function on the Shiny application
